---
title: "Basic 1"
author: "Eralda Gjika"
date: "9/15/2021"
output: html_document
---

# Examples
First importing some data from existing datasets in R (here I will use iris dataset)
```{r}
head(iris) # this will give us a view of the dataframe iris
str(iris)# this will give us the structure of the dataframe
```

# Example
```{r}
n = 50
x = 1:n
y = rnorm(n,10+2*x, 20)
z = rnorm(n, 5+3*x, 10)
Categors = rep(c("E","D","C","B","A"),each=n/5)
MoreCats = rep(c("a1","a2","a6","a4","a5","a6","a3","a8","a9","a2"),each=n/10)
Data2use = data.frame(Xthing=x,Ything=y, Zthing=z, Cats = Categors,MoreCats= MoreCats)
```

```{r}
plot(Data2use$Xthing, Data2use$Ything, type="p")
plot(Data2use$Xthing,Data2use$Ything,col=Data2use$Cats) #fails,how to fix?
plot(Data2use$Xthing,Data2use$Ything,col=factor(Data2use$Cats)) #works now
legend("topleft",c("A","B","C","D","E"),pch = 1,col = 1:5) #is this legend right?

```

# GGplot
```{r}
#install.packages("tidyverse")
library(tidyverse)
ggplot(Data2use, aes(x=Xthing,y=Ything))+geom_point()
ggplot(Data2use, aes(x=Xthing,y=Ything, colour = Cats))+
geom_point()
```



## References:
https://r4ds.had.co.nz/data-visualisation.html

The template:
ggplot(data = <DATA>) +
<GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))

Can be generalized to more overplayed plot types:

ggplot(data = <DATA>) +
<GEOM_FUNCTION1>(mapping = aes(<MAPPINGS1>))+
<GEOM_FUNCTION2>(mapping = aes(<MAPPINGS2>))

# Plot types
https://rstudio.com/wp-content/uploads/2015/03/ggplot2-cheatsheet.pdf

```{r}
ggplot(Data2use, aes(x=Xthing,y=Ything))+
geom_point()
ggplot(Data2use, aes( y=Ything ))+
geom_boxplot()
ggplot(Data2use, aes( y=Ything ))+
geom_density()
```
##Combine plot types
**Loess** = locally weighted regression:
Typically fits weighted regression over a very fine grid
```{r}
ggplot(Data2use, aes(x=Xthing,y=Ything))+
geom_point()+
geom_smooth()

ggplot(Data2use, aes(x=Xthing,y=Ything,colour=Cats))+
geom_point()+
geom_smooth(formula = y~x)
ggplot(Data2use, aes(x=Xthing,y=Ything))+
geom_point(aes(colour=Cats))+
geom_smooth(formula = y~x)
```
# Exercises from the textbook
• Hadley Wickham and Garrett Grolemund "R for Data Science", OReilly books:
https://r4ds.had.co.nz (online 3.2.4)
• Run ggplot(data = mpg). What do you see?
Answer: Just a plot window
```{r}
ggplot(data = mpg)
```
• How many rows are in mpg? How many columns?
```{r}
dim(mpg)# gives the dimession of the dataframe 234 rows (observations) and 11 columns (variables)
str(mpg)# gives the structure of the dataframe
```
• What does the drv variable describe? Read the help for ?mpg to find out.
Answer: 
drv variable is a character variable, 
drv -> the type of drive train, where f = front-wheel drive, r = rear wheel drive, 4 = 4wd
```{r}
?mpg
```

• Make a scatterplot of hwy vs cyl.
```{r}
ggplot(mpg,aes(x=cyl,y=hwy))+
         geom_point()
```

• What happens if you make a scatterplot of class vs drv? Why is the plot not
useful?
Answer: we can not clearly understand the number of observations for each combinations of class and drv categories (we have two categorical variables) 
```{r}
ggplot(mpg,aes(x=class,y=drv))+
         geom_point()
```
# Jittering
helps us to better view the situation when dealing with categorical variables

```{r}
ggplot(Data2use,aes(x=Cats,y=Ything))+
geom_point()+
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

ggplot(Data2use,aes(x=Cats,y=Ything))+
geom_jitter()+
theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```
# Multivariate data
```{r}
ggplot(Data2use, aes(x=Xthing,y=Ything, shape = Cats, colour = MoreCats))+
geom_point()

ggplot(Data2use, aes(x=Xthing,y=Ything, colour = Cats, alpha = Zthing))+
geom_point()
```

# Facets
Subplotting
• 
```{r}
ggplot(Data2use,aes(x=Xthing,y=Ything))+
geom_point()+
facet_wrap( ~Cats)

ggplot(Data2use,aes(x=Xthing,y=Ything))+
geom_point()+
facet_wrap( MoreCats~Cats)
```
